Synthesis of Linear Functional Programs

### 2021-07-13 article/article.tex
Memoization & run benchmarks

\mypara{SMT Solving} To typecheck and synth refinement types and dependent
functions, we need to contact an SMT solver to check satisfiability. For this,
we use the library \emph{sbv}~\cite{sbv}. Unfortunately, this library does not
support the usage of the SMT solvers to satisfy and output more complicated
first-order logic solutions -- since this is needed to synth refinement types
when right focused, we instead use unsafe library primitives to communicate with
the solver directly, constructing the logical formulas ourselves. For 
satisfying formulas, for instance, when typechecking, or left focused on a
refinement type, we use the library api successfully to get a veridct on
satisfiability.

\mypara{Memoization} ...

>>>>

\mypara{SMT Solving} To typecheck and synth refinement types and dependent
functions, we need to contact an SMT solver to check satisfiability. For this,
we use the library \emph{sbv}~\cite{sbv}. Unfortunately, this library does not
support the usage of the SMT solvers to satisfy and output more complicated
first-order logic uninterpreted functions -- since this is needed to synth refinement types
when right focused, we instead use unsafe library primitives to communicate with
the solver directly, constructing the logical formulas ourselves. For 
satisfying formulas, for instance, when typechecking, or left focused on a
refinement type, we use the library api successfully to get a veridct on
satisfiability.

\mypara{Memoization} Some proofs attempt to synth the same sub-goal with the
same premises more than once. Since these proof attempts might be expensive
performance-wise, and we can rely on the program's determinism (equal calls to
synth result in equal outcomes), we add to our synth memoization, an
optimization technique that stores function calls and returns cached values for
equal inputs. For our synth, two calls are equivalent when contexts $\Theta,
\Rho, \Gamma, \Delta$, $\Omega$, (or alternatively the left focus), and the
goal, are equivalent. Because comparing all of these is slow, our implementation
hashes the relevant values and uses the result as the key. Using LogicT, a
branch we want to memoize might fail -- it's desireable to also record that a
synthesis fails, as it happens more than successful results. We capture failure
of a logict computation using the $ifte$ function and represent failure or
success information with the Maybe type. We ran benchmarks to measure the impact
of memoization on the performance. It's more noticeable in synthesis with
multiple functions in the unrestricted context, or more complex
specifications:
%
\begin{center}
    \begin{tabular}{ |c|c|c|c|c| }
        \hline
        Goal & Time + $\sigma$ & TimeM + $\sigma$ \\
        \hline
        insertWithKey & 13.31 s ± 593.4 ms & 2.667 s ± 284.3 ms
    \end{tabular}
\end{center}
